---
title: "cis-QTL analysis and GWAS integration workflow"
author: "Anqi Wang and Gao Wang"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{cis-QTL analysis and GWAS integration workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


Here we demonstrate the use of the `pecotmr` package to perform
colocalization, TWAS and enrichment analysis between a pair of molecular QTL
and GWAS using de-identified example data shipped with the package.
All sample names, variant positions, and identifiers are synthetic.

## Set up your environment

Load the `pecotmr` package.

```{r load-pkgs}
library(pecotmr)
```

## Load data

```{r}
data(gwas_sumstats_example)
data(eqtl_region_example)
gwas <- gwas_sumstats_example
eqtl <- eqtl_region_example
```

```{r}
head(gwas)
```

```{r}
names(eqtl)
```

```{r}
head(eqtl$X[, 1:5])
```

```{r}
head(eqtl$y_res)
```

```{r}
dim(eqtl$X)
```

## Fine-mapping using SuSiE

```{r}
library(susieR)
library(tidyverse)
get_cs_index <- function(snps_idx, fitted_data) {
    idx <- tryCatch(
        which(
            pmap(list(a = fitted_data$sets$cs), function(a) snps_idx %in% a) %>% unlist()
        ),
        error = function(e) NA_integer_
    )
    if(length(idx) == 0) return(NA_integer_)
    return(idx)
}
signal_cutoff = 0.1
coverage = 0.7
susie_res <- susie(eqtl$X, eqtl$y_res,
                 L=20,
                 max_iter=500,
                 estimate_residual_variance=TRUE,
                 estimate_prior_variance=TRUE,
                 refine=TRUE,
                 compute_univariate_zscore=FALSE,
                 coverage=coverage)
```

```{r}
secondary_coverage = 0.2
susie_res$sets_secondary = susie_get_cs(susie_res, eqtl$X, coverage=secondary_coverage)
```

```{r}
compute_maf <- function(geno){
  f <- mean(geno,na.rm = TRUE)/2
  return(min(f, 1-f))
}
variants_index = c(which(susie_res$pip >= signal_cutoff), unlist(susie_res$sets$cs)) %>% unique %>% sort
maf = apply(eqtl$X, 2, compute_maf)[variants_index]
all_variants = colnames(eqtl$X)
variants = all_variants[variants_index]
pip = susie_res$pip[variants_index]
cs_info = map_int(variants_index, ~get_cs_index(.x, susie_res))
cs_index = ifelse(is.na(cs_info), 0, str_replace(names(susie_res$sets$cs)[cs_info], "L", "") %>% as.numeric)
univariate_res = univariate_regression(eqtl$X[, variants_index, drop=F], eqtl$y_res)
# here betahat and sebetahat were computed from standardized X and Y
susie_res$top_loci = cbind(variants, maf, univariate_res$betahat, univariate_res$sebetahat, pip, cs_index)
colnames(susie_res$top_loci) = c("variant_id", "maf", "bhat", "sbhat", "pip", "cs_index_primary")
rownames(susie_res$top_loci) = NULL
```

```{r}
susie_res$top_loci
```

Top PIP variants:

```{r}
sort(susie_res$pip, decreasing = TRUE)[1:5]
```

## TWAS

Use SuSiE weights,

```{r}
R = cor(eqtl$X)
```

```{r}
weights = susie_weights(susie_fit = susie_res)
```

```{r}
twas_z(weights, gwas$z, R=R)
```

```{r}
cor(eqtl$X%*%weights, eqtl$y_res)^2
```

Now select only the top loci,

```{r}
plot(weights)
```

```{r}
weights[!all_variants %in% susie_res$top_loci[,1]] = 0
```

```{r}
plot(weights)
```

```{r}
twas_z(weights, gwas$z, R=R)
```

Use elastic net (alpha=0.5),

```{r}
weights = glmnet_weights(eqtl$X, eqtl$y_res, alpha=0.5)
cor(eqtl$X%*%weights, eqtl$y_res)^2
twas_z(weights, gwas$z, R=R)
```

```{r}
plot(weights)
```

Use LASSO,

```{r}
weights = glmnet_weights(eqtl$X, eqtl$y_res, alpha=1)
cor(eqtl$X%*%weights, eqtl$y_res)^2
twas_z(weights, gwas$z, R=R)
```

```{r}
plot(weights)
```

Use mr.ash,

```{r}
lasso.weights = glmnet_weights(eqtl$X, eqtl$y_res, alpha=1)
weights = mrash_weights(eqtl$X, eqtl$y_res, init_prior_sd=TRUE, beta.init=lasso.weights)
cor(eqtl$X%*%weights, eqtl$y_res)^2
twas_z(weights, gwas$z, R=R)
```

```{r}
plot(weights)
```

```{r}
weights = mrash_weights(eqtl$X, eqtl$y_res, init_prior_sd=FALSE, beta.init=lasso.weights)
cor(eqtl$X%*%weights, eqtl$y_res)^2
twas_z(weights, gwas$z, R=R)
```

```{r}
plot(weights)
```

## Enrichment analysis

```{r}
data(gwas_finemapping_example)
data(qtl_finemapping_example)
gwas_path <- tempfile(fileext = ".rds")
qtl_path <- tempfile(fileext = ".rds")
saveRDS(gwas_finemapping_example, gwas_path)
saveRDS(qtl_finemapping_example, qtl_path)
xqtl_enrichment_wrapper(gwas_files = gwas_path, xqtl_files = qtl_path,
                        xqtl_finemapping_obj = c("context_1", "susie_result_trimmed"),
                        xqtl_varname_obj = c("context_1", "variant_names"))
```
